\chapter*{Abstract}
\addcontentsline{toc}{chapter}{Abstract}
As more and more businesses and organizations are moving away from having a their own dedicated server to renting server space in large server parks, like a cloud, security are becoming a bigger concern than ever before. When we want to perform computation on sensitive data stored on a cloud server a problem arises. As the cloud environment is considered untrusted, the data stored on the cloud server must be encrypted with our own encryption key to make sure the data is secure to attacks from both outside and inside the cloud. so, how can we efficiently perform a computation without exposing the data? One way to do this is to encrypt the sensitive data and store this on the cloud. To perform a computation the data would be downloaded to a private server and decrypted. The computation is performed and the data is encrypted and uploaded back to the cloud. The problem with this approach, it's  inefficient, especially when a lot of computation is done and it also requires a private server to perform the computation. Another approach is to use a Homomorphic Encryption(HE) scheme which enables us to perform computation on encrypted data without decrypting it first. This may sound like the holy grail, but there are some issues with HE like performance and key-size. Unfortunately Fully Homomorphic Encryption(FHE) libraries are very limited in Python. In this paper we have implemented a Python API based of a FHE C++ library called HElib and we will take a look at potential usages and limitations to the library. 
